Vulnerability Analysis of Container Images in Neuroimaging

### 2020-07-01 paper.tex

\author[1]{Bhupinder Kaur}
\author[1]{Aiman Hanna}
\author[1]{Tristan Glatard}

@@

Today, containers, which provide operating system level virtualization, are
popular. Correspondingly, they are widely used for scientific data analyses on 
High Performance Computing (HPC) clusters. However, the security of containers
still remains the concerning issue. Here, we focus on the security of container
images in particular. Container images are often built for the entire
duration of an analysis, so scientific researchers do not prefer to update
them. Consequently, their chances of containing vulnerabilities increases.  
In this study, we analyze vulnerabilities in container images used in 
HPC clusters.
We conduct two different
experiments, image update and image shrinking, to report the effect 
on the number of vulnerabilities.

@@

\textit{RQ3: Can the amount of vulnerabilities be reduced by shrinking the images?} 
Container images often include more software packages than required for a
typical analysis. We report on the impact of this additional software on
the presence of vulnerabilities.

@@

The next section describes the materials and tools used in the experiments,
including image scanners and container images. We then describe our image
update and shrinking methodology to reduce vulnerabilities. Results review
the detected vulnerabilities in container images, explain observed
discrepancies between scanners, and report on the efficiency of image
update and size reduction. As a conclusion, we provide a set of image
creation guidelines for a secure deployment on HPC clusters.

@@

Since the aim of this research work is to detect security vulnerabilities
in container images and attempt to examine the effect of different
manipulation of these images, it is needed to have a representative set
images. Additionally, for the purpose of detecting vulnerabilities,
utilization of image scanners is required. We utilize container
images from two popular application frameworks, as well as
four of the main image scanners.

@@

We use four popular image scanners, namely Anchore~\cite{github_2019},
Vuls~\cite{future-architect_2019}, Clair-scanner~\cite{arminc_2019} and
Stools~\cite{stools}. Anchore, Vuls and Clair-scanner support Docker
images while Stools supports Singularity. Different vulnerabilities may be
reported if scanning experiments take place on different dates.
To avoid such discrepancies, we freeze the vulnerability
databases used by these scanners as of 2019-09-25, when we initiate our
experiments. This allows for a fair comparison between the scanners.

@@

Anchore is an end-to-end, open-source container security platform. It
is a static
scanning tool that analyses container images and lists OS
packages, non-OS packages (Python, Java, Gem, and Npm), and files.
Anchore Engine can be used in two ways: through Docker image or Anchore
Command Line Interface (CLI).
It can be
installed directly on the host and used to inspect images.
In our experiments, we use Anchore Engine version 0.5.0 through Docker image \todo{insert name}, and
Anchore vulnerability database version 0.0.11.

Anchore compares package versions, which are present inside the
image, with the vulnerability database in order to flag vulnerabilities.
To get a vulnerability feed of various OS distributions, Anchore refers to
different sources, for instance, Ubuntu launchpad database to get
vulnerability data related to Ubuntu.

@@

Vuls is an open-source vulnerability scanner for Linux and FreeBSD. It
offers both static and dynamic scanning, and both local and remote
scanning. In our experiments, we use remote dynamic scanning, that is, Vuls
starts a container for every image to test, and queries the package manager in the image through SSH.
We use Vuls 0.9.0, executed through Docker image \todo{add Docker image name}. Vuls gets data from the Open Vulnerability and Assessment Language (OVAL) database.
It is worth noting that there is no OVAL data for Ubuntu 17.04 and 17.10 distributions,
meaning that images with these distributions cannot be scanned with Vuls.

\subsubsection{Clair-scanner}

Clair-scanner is a static vulnerability scanner for Docker containers.
It scans the image, prepares the list of
vulnerabilities, compares that list against a whitelist, and flags vulnerabilities
that are not present in the whitelist.
Clair-scanner utilizes the Ubuntu CVE tracker to obtain  Ubuntu vulnerability feed.
For that purpose, a Docker image is maintained that contains the Ubuntu CVE tracker database.
Whenever Ubuntu updates its database, this image is also updated with the required information.
We use Clair-scanner version 2.0.6, executed through  Docker image arminc/clair-local-scan:v2.0.6. 
We use another Docker image (arminc/clair-db:latest) for the vulnerability database, which was
last updated on 2019-09-18.

\subsubsection{Stools}

\todo{remove mentioning of CI, rewrite this paragraph.}
Continuous Integration (CI) allows smaller code changes, fault isolations,
faster release date, etc. Two types of CI are commonly used : TravisCI and
CircleCI. The former targets open source projects that need to be tested in
different environments, whereas the latter targets smaller projects where
integration needs to be done as fast as possible.

Stools uses CoreOS's Clair, which
is used for Docker images, in combination with CI to allow scanning of
Singularity images.
It involves spinning a Clair container during testing and building a
Singularity image, and scanning it before the image is finalized.

In our experiments, we use a Docker image (\href{https://hub.docker.com/r/vanessa/stools-clair}{vanessa/stools-clair:v3.2.1})
following the steps mentioned in \href{https://github.com/singularityhub/stools}{Stools Github} for scanning
Singularity images.

@@

Through our experiments, we find that image scanners report vulnerabilities 
differently depending on the different Linux release types. So, it is 
important to provide an overview of these Linux releases types. 
BIDS app and Boutiques images in our experiments have diverse Linux
distributions, such as Ubuntu, CentOS, Debian, and Alpine. Further, a particular
distribution has different releases in general. To illustrate, we here examine Ubuntu as an example,
since it is heavily utilized in our experiments. 
\todo{Explain the situation with centos, update figure 3 accordingly.}
However, other distributions also have
similar release trend. Ubuntu releases
can be of two types: Regular or Long Term Support (LTS).
Regular release is supported only for 9 months, whereas LTS is supported for 5 years. Once LTS support
reaches End-Of-Life (EOL), Ubuntu decides whether to provide extended security support or not, which comes
in the form of Extended Security Maintenance (ESM) release.
ESM is a paid service that can be requested from Ubuntu to get security 
updates even after the EOL of a particular Ubuntu release. ESM and LTS releases of a particular Ubuntu distribution have different vulnerability
databases, because after EOL of a release, a vulnerability is fixed only in packages of ESM; however, it still
exists in LTS.

@@

\section{Our Approach}

In this section, we present our approach, where we attempt the reduction
of security vulnerabilities in images by updating or
shrinking these images.

@@

If vulnerabilities exist in the container images, then our first approach
is to update these images in order to see the effects of such updates
on the vulnerabilities.
Updating an image involves updating all software packages that are
inside that image. We use a 
\href{https://github.com/kaurbhupinder/Vulnerability-Analysis/blob/master/Scripts/update/update.sh}{bash script} 
to update all images. That script checks the
package manager, which is present inside the image, and updates the image
accordingly. However, this action affects the reproducibility
of that image. For example, if a package is used by a scientist to conduct an experiment
and later that package is updated, other scientists want to reproduce initial results may actually
get different results due to updates.
%which means that the image cannot be used to verify the research
%findings of other scientists.

\subsection{Image Reduction/Minimization Process}
On the other hand, if it is desired to reduce the number of vulnerabilities without affecting the reproducibility of the image,
then we revert to image size reduction instead of image update. An interesting choice to do so would be to use 
Neurodocker~\cite{neurodocker}.
This tool traces and deletes unnecessary files.
However, it does not affect the number of installed packages.
Since image scanners report vulnerabilities according to the installed
packages. Therefore, after
minimization the same number of vulnerabilities is still reported.
Consequently, we need to reduce the number of installed packages
in the images by deleting unnecessary packages. For that, it is required to figure
out which packages we need for the pipeline to function properly. To allow for that, we utilize
Reprozip~\cite{rampin2016reprozip}, which is a tool that
is used to make applications reproducible. This tool traces operating system calls used by the
application while execution, and then based on this trace, it recognizes binaries, files, environment variables,
and library dependencies that are used by the application for future re-execution. If the OS is
Debian or RPM-based, Reprozip also
uses an available package manager to acquire a list of distribution packages from where these files come from.
So, a list of packages that are needed for the application re-execution is determined. Following that, the list
is expanded by adding important and required packages that are essential
for the system to function. Finally, we figure out all the dependencies of these packages using Debtree
~\cite{debtree}.
Effectively all needed packages are kept and remaining unnecessary packages are removed. We use
\href{https://github.com/kaurbhupinder/Vulnerability-Analysis/tree/master/Scripts/minification}{bash scripts} for
the image size reduction process, which install Reprozip in the image, collect Reprozip trace, and
finally delete all unnecessary packages.

@@

Figure~\ref{fig:venn} represents the number of vulnerabilities reported by
the scanners and the overlap between them.

>>>>

\author[1]{Bhupinder Kaur}
\author[1]{Mathieu Dugr\'e}
\author[1]{Aiman Hanna}
\author[1]{Tristan Glatard}

@@

Today, containers, which provide operating system level virtualization, are
popular. Correspondingly, they are widely used for scientific data analyses on 
High Performance Computing (HPC) clusters. However, the security of containers
still remains the concerning issue. Here, we focus on the security of container
images in particular. Container images are often built for the entire
duration of an analysis, so scientific researchers do not prefer to update
them. Consequently, their chances of containing vulnerabilities increases.  
In this study, we analyze vulnerabilities in container images used in 
HPC clusters.
We conduct two different
experiments, image update and image minimization, to report the effect 
on the number of vulnerabilities.

@@

\textit{RQ3: Can the amount of vulnerabilities be reduced by minimizing the images?} 
Container images often include more software packages than required for a
typical analysis. We report on the impact of this additional software on
the presence of vulnerabilities.

@@

The next section describes the materials and tools used in the experiments,
including image scanners and container images. We then describe our image
update and minimization methodology to reduce vulnerabilities. Results review
the detected vulnerabilities in container images, explain observed
discrepancies between scanners, and report on the efficiency of image
update and size reduction. As a conclusion, we provide a set of image
creation guidelines for a secure deployment on HPC clusters.

@@

Since the aim of this research work is to detect security vulnerabilities
in container images and attempt to examine the effect of different
manipulation of these images, it is needed to have a representative set of
images. Additionally, for the purpose of detecting vulnerabilities,
utilization of image scanners is required. We utilize container
images from two popular application frameworks, as well as
four of the main image scanners.

@@

We use four popular image scanners, namely Anchore~\cite{github_2019},
Vuls~\cite{future-architect_2019}, Clair~\cite{arminc_2019} and
Singularity Container Tools --- Stools~\cite{stools}. Anchore, Vuls and Clair support Docker
images while Stools support Singularity. Different vulnerabilities may be
reported if scanning experiments take place on different dates.
To avoid such discrepancies, we freeze the vulnerability
databases used by these scanners as of 2019-09-25, when we initiate our
experiments. This allows for a fair comparison between the scanners.

@@

Anchore is an end-to-end, open-source container security platform. It
analyses container images and lists OS
packages, non-OS packages (Python, Java, Gem, and Npm), and files.
In our experiments, we use Anchore Engine version 0.5.0 through Docker image \todo{insert name}, and
Anchore vulnerability database version 0.0.11.

Anchore compares package versions, which are present inside the
image, with the vulnerability database in order to flag vulnerabilities.
To get a vulnerability feed of various OS distributions, Anchore refers to
different sources, for instance, Ubuntu launchpad database to get
vulnerability data related to Ubuntu \tristan{You need to be accurate on these sources, this is not ``for instance". 
You should list all the sources used in the experiments, for all image types: Ubuntu but also CentOS, Debian, Alpine.}.

@@

Vuls is an open-source vulnerability scanner for Linux and FreeBSD. It
offers both static and dynamic scanning, and both local and remote
scanning. In our experiments, we use remote dynamic scanning, that is, Vuls
starts a container for every image to test, and queries the package manager
in the image through SSH. We use Vuls 0.9.0, executed through Docker image
\todo{add Docker image name}. Vuls gets data from the Open Vulnerability
and Assessment Language (OVAL) database \tristan{Does it contain
vulnerabilities for all distributions?} \tristan{The Vuls website says that it uses multiple vulnerability databases: https://vuls.io/docs/en/main-features.html Are you sure 
that you only used OVAL and if so, why?}. It is worth noting that there is
no OVAL data for Ubuntu 17.04 and 17.10 distributions, meaning that images
with these distributions cannot be scanned with Vuls.

\subsubsection{Clair}

Clair (\url{https://github.com/quay/clair}) is an open-source and extensible
vulnerability scanner for Docker and appc container images, developed by
CoreOS (now Container Linux), a Linux distribution to deploy container
clusters. Clair has a client-server architecture, in which the server
maintains a database of vulnerabilities. We use Clair through Clair scanner
(\url{https://github.com/arminc/clair-scanner}), a tool to facilitate the testing
of container images against a local Clair server. Clair scanner scans the image,
prepares the list of vulnerabilities, compares that list against a
whitelist, and flags vulnerabilities that are not present in the whitelist
\tristan{What is this whitelist about, and which one did you use?}. Clair
scanner maintains a Docker image with the up-to-date Ubuntu CVE tracker
database, used for the analysis \tristan{Please mention the vulnerability sources for all distributions}.
 We use Clair-scanner version 2.0.6, executed through
Docker image \texttt{arminc/clair-local-scan:v2.0.6}. For the vulnerability
database, we use Docker image \texttt{arminc/clair-db:latest}, last
updated on 2019-09-18.

\subsubsection{Singularity Tools}

Singularity Tools (Stools)
are an extension of Clair scanner for Singularity images. Stools
directly scan the image files in a local installation of the Clair
scanner, to circumvent the Docker-specific requirements in the Clair API.
In our experiments, we use Singularity Tools version 3.2.1  through Docker
image
\texttt{vanessa/stools-clair:v3.2.1}
and we follow the steps mentioned in
\href{https://github.com/singularityhub/stools}{Stools Github} to scan
Singularity images. \tristan{which docker images did you use for clair and clair database?}

@@

\tristan{This section is still very confusing. I can rewrite it but would first need to
have a clear picture of vulnerability statuses across the tested Linux distributions. 
Are vulnerability statuses present only in Ubuntu, or are there statuses in other distributions too?
If so, what are the vulnerability statuses in the other distributions?
}
Through our experiments, we find that image scanners report vulnerabilities 
differently depending on the different Linux release types. So, it is 
important to provide an overview of these Linux releases types. 
BIDS app and Boutiques images in our experiments have diverse Linux
distributions, such as Ubuntu, CentOS, Debian, and Alpine. Further, a particular
distribution has different releases in general. To illustrate, we here examine Ubuntu as an example,
since it is heavily utilized in our experiments. 
\todo{Explain the situation with centos, update figure 3 accordingly.}
However, other distributions also have
similar release trend. Ubuntu releases
can be of two types: Regular or Long Term Support (LTS).
Regular release is supported only for 9 months, whereas LTS is supported for 5 years. Once LTS support
reaches End-Of-Life (EOL), Ubuntu decides whether to provide extended security support or not, which comes
in the form of Extended Security Maintenance (ESM) release.
ESM is a paid service that can be requested from Ubuntu to get security 
updates even after the EOL of a particular Ubuntu release. ESM and LTS releases of a particular Ubuntu distribution have different vulnerability
databases, because after EOL of a release, a vulnerability is fixed only in packages of ESM; however, it still
exists in LTS.

@@

\section{Reducing the amount of vulnerabilities}

In this section, we present our approach to reduce the number of
vulnerabilities in container images  by updating and minimizing these
images.

@@

If vulnerabilities exist in the container images, then our first approach
is to update these images in order to see the effects of such updates
on the vulnerabilities.
Updating an image involves updating all software packages that are
inside that image. We use a 
\href{https://github.com/kaurbhupinder/Vulnerability-Analysis/blob/master/Scripts/update/update.sh}{bash script} 
to update all images. That script checks the
package manager, which is present inside the image, and updates the image
accordingly. However, this action affects the reproducibility
of that image. For example, if a package is used in an experiment
and later that package is updated, subsequent instances of the experiment may actually
get different results.
%which means that the image cannot be used to verify the research
%findings of other scientists.

\subsection{Image Minimization Process}
We revert to image minimization to reduce the number of vulnerabilities without affecting results reproducibility.
An interesting choice to do so would be to use 
Neurodocker~\cite{neurodocker}.
This tool traces and deletes unnecessary files.
However, it does not affect the number of installed packages.
Since image scanners report vulnerabilities according to the installed
packages. Therefore, after
minimization the same number of vulnerabilities is still reported.
Consequently, we need to reduce the number of installed packages
in the images by deleting unnecessary packages. For that, it is required to figure
out which packages we need for the pipeline to function properly. To allow for that, we utilize
Reprozip~\cite{rampin2016reprozip}, which is a tool that
is used to make applications reproducible \tristan{No it doesn't make applications reproducible}. This tool traces operating system calls used by the
application while execution, and then based on this trace, it recognizes binaries, files, environment variables,
and library dependencies that are used by the application for future re-execution. If the OS is
Debian or RPM-based, Reprozip also
uses an available package manager to acquire a list of distribution packages from where these files come from.
So, a list of packages that are needed for the application re-execution is determined. \tristan{How did you run the applications? Which command lines, why, etc} Following that, the list
is expanded by adding important and required packages that are essential
for the system to function. Finally, we figure out all the dependencies of these packages using Debtree
~\cite{debtree} \tristan{does debtree work in centos? if not, how do you minimize centos images?}.
Effectively all needed packages are kept and remaining unnecessary packages are removed. We use
\href{https://github.com/kaurbhupinder/Vulnerability-Analysis/tree/master/Scripts/minification}{bash scripts} for
the image size reduction process \tristan{Can we please have at least a README for these scripts?}, which install Reprozip in the image, collect Reprozip trace, and
finally delete all unnecessary packages.

@@

Figure~\ref{fig:venn} represents the number of vulnerabilities reported by
the scanners and the overlap between them. \tristan{According to our discussion, this is only for Ubuntu. This should be clarified.}
